---
title: "README"
author: "Linoja"
date: "`r Sys.Date()`"
output: github_document
---

# countmissingvalues

## Package Introduction and Description
The `countmissingvalues` package provides a function designed to help users count missing values within a dataset by specified groups. This function is particularly useful for gaining insights into the distribution of missing data across different groups, helping to facilitate data cleaning and preprocessing tasks.

### Key Functionality
The main function, `countmissingvalues()`, takes a dataset, a grouping column, and an optional parameter to control the structure of the output. The function groups data based on the specified column and calculates the count of missing values for all other columns, providing an organized view of missing data patterns.

## Installation Instructions
To install the `countmissingvalues` package from GitHub, you’ll need to use the `devtools` package. Here’s how to do it:
```{r}
# Install devtools if it's not already installed
install.packages("devtools")

# Install countmissingvalues package from GitHub
devtools::install_github("stat545ubc-2024/countmissingvalues")
```

### Demonstrated Usage
Once installed, you can start using the countmissingvalues() function as shown below. The function’s parameters include:

data: A data frame or tibble containing the dataset.
group_col: The column to group by when counting missing values.
.groups (optional): Controls grouping in the output, defaulting to "drop".


### Example Dataset
For demonstration, you can use the R's built-in airquality dataset, which contains air quality measurements in New York for several months. Since airquality is built into R, users don’t need to upload or install additional datasets to run the examples below.

Examples
1. Basic Usage
Count missing values in the airquality dataset, grouped by Month:

```{r}
library(countmissingvalues)
count_all_missing_by_group(airquality, Month)
```

2. Using the Pipe Syntax
The function supports piping, allowing you to pass the dataset directly into the function:

```{r}
airquality |> countmissingvalues::count_all_missing_by_group(Month)
```

3. Keeping Grouped Output
To retain the grouping in the output tibble, set .groups = "keep":

```{r}
count_all_missing_by_group(airquality, Month, .groups = "keep")
```


### License
This package is licensed under the MIT License.
